Carbonate chemistry at Ny-Ålesund

### 2021-04-13 awipev-CO2.Rmd
back to a functional version

```{r read observatory data, cleanup and prepare ferrybox data, echo=FALSE, warning=FALSE, message=FALSE}
d_hour <-
  readRDS(file = paste0(
    path,
    "fb_awipev-co2_server/ny-alesund/data/NRT_data/nydata_hour.rds"
  ))
d_hour <- d_hour %>%
  dplyr::filter(datetime <= as.POSIXct("2021-01-10")) %>% #cutoff date
  dplyr::select(-pco2_inst) %>% # this is the serial number of pCO2 sensor, not needed
  dplyr::rename(
    pco2_fb = pco2_corr_filtered,
    temp_dur = temp_dur_filtered,
    pco2_raw = pco2_raw_filtered,
    sal_fb = sal_fb_filtered,
    temp_insitu_11m = temp_insitu_11m_filtered,
    temp_fb = temp_fb_filtered,
    ph_dur = ph_dur_filtered
  ) %>% 
  arrange(datetime)
# read a file provided by Philipp on 2021-04-09 which fills most gaps of pressure data
# depth_complete replaces all other depth and pressure variables
svl_remos_depth <- tibble(
  readRDS(file = paste0(
    path,
    "fb_data/svl_remos_depth.rds"
  )))
d_hour <- dplyr::left_join(d_hour, svl_remos_depth, by = 'datetime') %>% 
  dplyr::select(-adcp_fill, -ctd_fill) %>% 
  dplyr::rename(depth = depth_complete)
# check_plot <- d_hour %>% 
#   ggplot() + geom_point(aes(x=datetime, y=depth, col='blue', alpha = 0.7)) +
#   geom_point(aes(x=datetime, y=depth, col='red', alpha = 0.3))
d_all <- as_tibble(d_hour) #d_all will be the merger
d_all <- d_all %>%
  dplyr::mutate_all(function(x) ifelse(is.nan(x), NA, x)) %>% # replace all NaN by NA
  dplyr::mutate(depth_group = ifelse( # create new variable depth
      depth <= 2,
      1,
      ifelse(
        depth > 2 & depth <= 4,
        3,
        ifelse(
          depth > 4 & depth <= 6,
          5,
          ifelse(
            depth > 6 & depth <= 8,
            7,
            ifelse(depth > 8, 10, NA)
          )
        )
      )
    )) %>%
  dplyr::rowwise() %>% 
  dplyr::mutate(n_depth_ctd = sum(is.na( # create new variable to identify the hours for which there are CTD data at more than 1 depth 
  c(
    sal_insitu_ctd_1m,
    sal_insitu_ctd_3m,
    sal_insitu_ctd_5m,
    sal_insitu_ctd_7m,
    sal_insitu_ctd_9m
  )
), na.rm = TRUE)) %>% 
  ungroup()
#restore datetime, date which are mangled by the mutates above, arrange ascending
d_all <- d_all %>% 
  dplyr::mutate(datetime = d_hour$datetime, 
                date = d_hour$date,
                year = lubridate::year(d_hour$datetime), # year as a number
                monthd = lubridate::month(d_hour$datetime), # month as a number
                montht = lubridate::month(d_hour$datetime, label = TRUE, abbr = TRUE) # abbreviated months
  )
rm(d_hour)
```

@@

# add sal_fb + sal_insitu_ctd + temp_fb + temp_insitu_ctd from d_all into d, to use pHinsi
# use the closest time to add them
closest <- which.closest(d$datetime, d_all$datetime) # closest date and time
d$closest_datetime <- d_all$datetime[closest]
# if the closest value is more than 30 min away, then we do not take it
d$closest_datetime[abs(d$datetime - d_all$datetime[closest]) > 30*60] <- NA
# create closest_datetime in d_all (copy of datetime)
d_all$closest_datetime <- d_all$datetime
d <- left_join(d, d_all %>%
                 dplyr::select(
                   c(closest_datetime,
                     sal_fb,
                     temp_fb,
                     temp_mix,
                     sal_mix,
                     temp_insitu_11m,
                     temp_insitu_ctd,
                     depth
                   )
                 ), by = "closest_datetime") %>% 
  dplyr::select(-closest_datetime)

# Conversion of spec pH of in situ samples to in situ temp and 4 °C
pH_seafet <- d %>%
  dplyr::filter(!is.na(ph_s_sf) & !is.na(temp_insitu_ctd))
pH_seafet <- pH_seafet %>% 
  dplyr::mutate(
    ph_s_sf_temp_insi = pHinsi(pH = pH_seafet$ph_s_sf, 
                         ALK = pH_seafet$at_mean_interp*1e-6, Tinsi = pH_seafet$temp_insitu_ctd, 
                         Tlab = pH_seafet$temp_s_lab, Pinsi = pH_seafet$depth/10, 
                         S = pH_seafet$sal_mix, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
                         pHscale = "T", b="u74"),
   ph_s_sf_temp_4 = pHinsi(pH = pH_seafet$ph_s_sf, 
                         ALK = pH_seafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_seafet$temp_s_lab, Pinsi = pH_seafet$depth/10, 
                         S = pH_seafet$sal_mix, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
                         pHscale = "T", b="u74")
  )
# Conversion of spec pH of FerryBox samples to FerryBox temp
pH_durafet <- d %>%
  tidyr::drop_na(ph_s_dur, sal_fb, temp_fb)
pH_durafet <- pH_durafet %>% 
  dplyr::mutate(
    ph_s_dur_temp_fb = pHinsi(pH = pH_durafet$ph_s_dur, 
                              ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = pH_durafet$temp_fb, 
                              Tlab = pH_durafet$temp_s_lab, Pinsi = 0, S = pH_durafet$sal_fb, 
                              Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", pHscale = "T", b="u74"),
   ph_s_dur_temp_4_l = pHinsi(pH = pH_durafet$ph_s_dur, # lueker
                         ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_durafet$temp_s_lab, Pinsi = 0, 
                         S = pH_durafet$sal_fb, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
                         pHscale = "T", b="u74"),
   ph_s_dur_temp_4_s20 = pHinsi(pH = pH_durafet$ph_s_dur, # Sulpis
                         ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_durafet$temp_s_lab, Pinsi = 0, 
                         S = pH_durafet$sal_fb, Pt = 0, Sit = 0, k1k2 = "s20", kf = "dg", ks="d", 
                         pHscale = "T", b="u74"),       # Millero 02
   ph_s_dur_temp_4_p18 = pHinsi(pH = pH_durafet$ph_s_dur, # Papadimitriou
                         ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_durafet$temp_s_lab, Pinsi = 0, 
                         S = pH_durafet$sal_fb, Pt = 0, Sit = 0, k1k2 = "p18", kf = "dg", ks="d", 
                         pHscale = "T", b="u74"),
   ph_s_dur_temp_4_m02 = pHinsi(pH = pH_durafet$ph_s_dur,  # Millero 02
                         ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_durafet$temp_s_lab, Pinsi = 0, 
                         S = pH_durafet$sal_fb, Pt = 0, Sit = 0, k1k2 = "m02", kf = "dg", ks="d", 
                         pHscale = "T", b="u74")

@@

# pH seaFET Final
# add despike with n=2 k = 5762 to the final ph corrected data
# despike me semble inutile, faire plutôt un simple des outliers qui, vérification faite, n'est même pas nécessaire
# Il faut par contre créér une nouvelle variable avec pH deeper than 8 m, called *_9m for consistency with salinity and temp
d_all <- d_all %>%
  #dplyr::mutate(pHint_tot_sf = despike(d_all$pHint_tot_sf, reference= "median", n=2, k=5761, replace="NA")) %>%
  dplyr::mutate(pHint_tot_sf = seafet_calib$pHint_tot_sf) %>%
  dplyr::mutate(pHint_tot_sf_1m = ifelse(depth > 0 & depth < 2, pHint_tot_sf, NA),
                pHint_tot_sf_3m = ifelse(depth >= 2 & depth < 4, pHint_tot_sf, NA),
                pHint_tot_sf_5m = ifelse(depth >= 4 & depth < 6, pHint_tot_sf, NA),
                pHint_tot_sf_7m = ifelse(depth >= 6 & depth < 8, pHint_tot_sf, NA),
                pHint_tot_sf_9m = ifelse(depth >= 8 , pHint_tot_sf, NA),
                depth_discrete = case_when(
      !is.na(pHint_tot_sf_1m) ~ 1,
      !is.na(pHint_tot_sf_3m) ~ 3,
      !is.na(pHint_tot_sf_5m) ~ 5,
      !is.na(pHint_tot_sf_7m) ~ 7,
      !is.na(pHint_tot_sf_9m) ~ 9
      ),
      depth_discrete = as.factor(depth_discrete)
  )

>>>>

```{r read observatory data, cleanup and prepare ferrybox data, echo=FALSE, warning=FALSE, message=FALSE}
d_hour <-
  readRDS(file = paste0(
    path,
    "fb_awipev-co2_server/ny-alesund/data/NRT_data/nydata_hour.rds"
  ))
d_hour <- d_hour %>%
  dplyr::filter(datetime <= as.POSIXct("2021-01-10")) %>% #cutoff date
  dplyr::select(-pco2_inst) %>% # this is the serial number of pCO2 sensor, not needed
  dplyr::rename(
    pco2_fb = pco2_corr_filtered,
    temp_dur = temp_dur_filtered,
    pco2_raw = pco2_raw_filtered,
    sal_fb = sal_fb_filtered,
    temp_insitu_11m = temp_insitu_11m_filtered,
    temp_fb = temp_fb_filtered,
    ph_dur = ph_dur_filtered
  ) %>% 
  arrange(datetime)
# read a file provided by Philipp on 2021-04-09 which fills most gaps of pressure data
# depth_complete replaces all other depth and pressure variables
svl_remos_depth <- tibble(
  readRDS(file = paste0(
    path,
    "fb_data/svl_remos_depth.rds"
  )))
d_hour <- dplyr::left_join(d_hour, svl_remos_depth, by = 'datetime') %>% 
  dplyr::select(-adcp_fill, -ctd_fill) %>% 
  dplyr::rename(depth = depth_complete)
# check_plot <- d_hour %>% 
#   ggplot() + geom_point(aes(x=datetime, y=depth, col='blue', alpha = 0.7)) +
#   geom_point(aes(x=datetime, y=depth, col='red', alpha = 0.3))
d_all <- as_tibble(d_hour) #d_all will be the merger
d_all <- d_all %>%
  dplyr::mutate_all(function(x) ifelse(is.nan(x), NA, x)) %>% # replace all NaN by NA
  dplyr::mutate(depth_group = ifelse( # create new variable depth
      pressure_insitu_ctd <= 2,
      1,
      ifelse(
        pressure_insitu_ctd > 2 & depth <= 4,
        3,
        ifelse(
          pressure_insitu_ctd > 4 & depth <= 6,
          5,
          ifelse(
            pressure_insitu_ctd > 6 & depth <= 8,
            7,
            ifelse(pressure_insitu_ctd > 8, 10, NA)
          )
        )
      )
    )) %>%
  dplyr::rowwise() %>% 
  dplyr::mutate(n_depth_ctd = sum(is.na( # create new variable to identify the hours for which there are CTD data at more than 1 depth 
  c(
    sal_insitu_ctd_1m,
    sal_insitu_ctd_3m,
    sal_insitu_ctd_5m,
    sal_insitu_ctd_7m,
    sal_insitu_ctd_9m
  )
), na.rm = TRUE)) %>% 
  ungroup()
#restore datetime, date which are mangled by the mutates above, arrange ascending
d_all <- d_all %>% 
  dplyr::mutate(datetime = d_hour$datetime, 
                date = d_hour$date,
                year = lubridate::year(d_hour$datetime), # year as a number
                monthd = lubridate::month(d_hour$datetime), # month as a number
                montht = lubridate::month(d_hour$datetime, label = TRUE, abbr = TRUE) # abbreviated months
  )
rm(d_hour)
```

@@

# add sal_fb + sal_insitu_ctd + temp_fb + temp_insitu_ctd from d_all into d, to use pHinsi
# use the closest time to add them
closest <- which.closest(d$datetime, d_all$datetime) # closest date and time
d$closest_datetime <- d_all$datetime[closest]
# if the closest value is more than 30 min away, then we do not take it
d$closest_datetime[abs(d$datetime - d_all$datetime[closest]) > 30*60] <- NA
# create closest_datetime in d_all (copy of datetime)
d_all$closest_datetime <- d_all$datetime
d <- left_join(d, d_all %>%
                 dplyr::select(
                   c(closest_datetime,
                     sal_fb,
                     temp_fb,
                     temp_mix,
                     sal_mix,
                     temp_insitu_11m,
                     temp_insitu_ctd,
                     pressure_insitu_ctd
                   )
                 ), by = "closest_datetime") %>% 
  dplyr::select(-closest_datetime)

# Conversion of spec pH of in situ samples to in situ temp and 4 °C
pH_seafet <- d %>%
  dplyr::filter(!is.na(ph_s_sf) & !is.na(temp_insitu_ctd))
pH_seafet <- pH_seafet %>% 
  dplyr::mutate(
    ph_s_sf_temp_insi = pHinsi(pH = pH_seafet$ph_s_sf, 
                         ALK = pH_seafet$at_mean_interp*1e-6, Tinsi = pH_seafet$temp_insitu_ctd, 
                         Tlab = pH_seafet$temp_s_lab, Pinsi = pH_seafet$pressure_insitu_ctd/10, 
                         S = pH_seafet$sal_mix, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
                         pHscale = "T", b="u74"),
   ph_s_sf_temp_4 = pHinsi(pH = pH_seafet$ph_s_sf, 
                         ALK = pH_seafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_seafet$temp_s_lab, Pinsi = pH_seafet$pressure_insitu_ctd/10, 
                         S = pH_seafet$sal_mix, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
                         pHscale = "T", b="u74")
  )
# Conversion of spec pH of FerryBox samples to FerryBox temp
pH_durafet <- d %>%
  tidyr::drop_na(ph_s_dur, sal_fb, temp_fb)
pH_durafet <- pH_durafet %>% 
  dplyr::mutate(
    ph_s_dur_temp_fb = pHinsi(pH = pH_durafet$ph_s_dur, 
                              ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = pH_durafet$temp_fb, 
                              Tlab = pH_durafet$temp_s_lab, Pinsi = 0, S = pH_durafet$sal_fb, 
                              Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", pHscale = "T", b="u74"),
   ph_s_dur_temp_4_l = pHinsi(pH = pH_durafet$ph_s_dur, # lueker
                         ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_durafet$temp_s_lab, Pinsi = 0, 
                         S = pH_durafet$sal_fb, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
                         pHscale = "T", b="u74"),
   ph_s_dur_temp_4_s20 = pHinsi(pH = pH_durafet$ph_s_dur, # Sulpis
                         ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_durafet$temp_s_lab, Pinsi = 0, 
                         S = pH_durafet$sal_fb, Pt = 0, Sit = 0, k1k2 = "s20", kf = "dg", ks="d", 
                         pHscale = "T", b="u74"),       # Millero 02
   ph_s_dur_temp_4_p18 = pHinsi(pH = pH_durafet$ph_s_dur, # Papadimitriou
                         ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_durafet$temp_s_lab, Pinsi = 0, 
                         S = pH_durafet$sal_fb, Pt = 0, Sit = 0, k1k2 = "p18", kf = "dg", ks="d", 
                         pHscale = "T", b="u74"),
   ph_s_dur_temp_4_m02 = pHinsi(pH = pH_durafet$ph_s_dur,  # Millero 02
                         ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_durafet$temp_s_lab, Pinsi = 0, 
                         S = pH_durafet$sal_fb, Pt = 0, Sit = 0, k1k2 = "m02", kf = "dg", ks="d", 
                         pHscale = "T", b="u74")

@@

# pH seaFET Final
# add despike with n=2 k = 5762 to the final ph corrected data
# despike me semble inutile, faire plutôt un simple des outliers qui, vérification faite, n'est même pas nécessaire
# Il faut par contre créér une nouvelle variable avec pH deeper than 8 m, called *_9m for consistency with salinity and temp
d_all <- d_all %>%
  #dplyr::mutate(pHint_tot_sf = despike(d_all$pHint_tot_sf, reference= "median", n=2, k=5761, replace="NA")) %>%
  dplyr::mutate(pHint_tot_sf = seafet_calib$pHint_tot_sf) %>%
  dplyr::mutate(pHint_tot_sf_1m = ifelse(pressure_insitu_ctd > 0 & pressure_insitu_ctd < 2, pHint_tot_sf, NA),
                pHint_tot_sf_3m = ifelse(pressure_insitu_ctd >= 2 & pressure_insitu_ctd < 4, pHint_tot_sf, NA),
                pHint_tot_sf_5m = ifelse(pressure_insitu_ctd >= 4 & pressure_insitu_ctd < 6, pHint_tot_sf, NA),
                pHint_tot_sf_7m = ifelse(pressure_insitu_ctd >= 6 & pressure_insitu_ctd < 8, pHint_tot_sf, NA),
                pHint_tot_sf_9m = ifelse(pressure_insitu_ctd >= 8 , pHint_tot_sf, NA),
                depth_discrete = case_when(
      !is.na(pHint_tot_sf_1m) ~ 1,
      !is.na(pHint_tot_sf_3m) ~ 3,
      !is.na(pHint_tot_sf_5m) ~ 5,
      !is.na(pHint_tot_sf_7m) ~ 7,
      !is.na(pHint_tot_sf_9m) ~ 9
      ),
      depth_discrete = as.factor(depth_discrete)
  )
